Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-07-25 experiments.tex
Added ACM style tables

As regards the class of benchmark being considered, a set applications representative of some common
frequently running code kernels has been selected. Table~\ref{tab:bench} shows the set of applications
along with a brief description.
\begin{table}
	\centering
	\caption{Benchmarks}
	\label{tab:bench}
	\begin{tabular}{ll}
	\hline
	\multicolumn{1}{c}{Benchmark} & \multicolumn{1}{c}{Application} \\
	\hline
	Alloca\_test & Memory array allocation test \\
	Bmm & Matrices multiplication and elements sum \\
	Fir\_int & Finite impulse response \\
	Mm & Floating point matrices multiplication \\
	Sha256 & Cryptocurrency header Hashing \\
	Wave & Wavefront computation \\
	\hline
	\end{tabular}
\end{table}

\subsection{Evaluation Metrics}

@@

\begin{table*}
  \centering
  \begin{tabular}{ccc}
    \includegraphics[width=0.30\textwidth]{pictures/alloca_100.eps} &
    \includegraphics[width=0.30\textwidth]{pictures/bmm_100.eps} & 
    \includegraphics[width=0.30\textwidth]{pictures/fir_int100.eps} \\
    \includegraphics[width=0.30\textwidth]{pictures/mm_100.eps} &
    \includegraphics[width=0.30\textwidth]{pictures/sha_100.eps} &
    \includegraphics[width=0.30\textwidth]{pictures/wave_100.eps} 
  \end{tabular}
  \caption{Pareto fronts found by PS and GA for a fixed budget of 100 configurations.}
  \label{fig:pareto_fronts_100}
\end{table*}

\begin{table*}
  \centering
  \begin{tabular}{ccc}
    \includegraphics[width=0.30\textwidth]{pictures/alloca_500.eps} &
    \includegraphics[width=0.30\textwidth]{pictures/bmm_500.eps} & 
    \includegraphics[width=0.30\textwidth]{pictures/fir_int500.eps} \\
    \includegraphics[width=0.30\textwidth]{pictures/mm_500.eps} &
    \includegraphics[width=0.30\textwidth]{pictures/sha_500.eps} &
    \includegraphics[width=0.30\textwidth]{pictures/wave_500.eps} 
  \end{tabular}
  \caption{Pareto fronts found by PS and GA for a fixed budget of 500 configurations.}
  \label{fig:pareto_fronts_500}
\end{table*}

Low budget Pareto sets are shown in
Figure~\ref{fig:pareto_fronts_100}. With a so small number of
configuration being investigated the two sets are often overlapping,
being also strongly depended upon the application being considered.
These results are obtained in about 5 eras, and the main aim was
demonstrating that with a so limited budget PS still can compete with
genetic approach while even discovering very different Pareto fronts.
Continuing with a more extended simulation budget as depicted in
Figure~\ref{fig:pareto_fronts_500}, we can qualitatively observe that
even if PS does not strictly outperforms genetic in terms of dominance, it shows in
some cases largely Pareto in terms of extension. We can intuitively
relate this to the ``novelty-based'' score system, so that instead of
focusing on getting better points inside a given range it tries to
enlarge the range itself, resulting in less dense but more extended
Paretos. Of course we cannot classify this different behaviour as
``better'' or ``worse'' as compared to the genetic one, it depends on
various design factors such as desired results granularity, prefixed objective
range constraints, estimated error of the models used etc~\ldots.

In order to evaluate the Paretos from a quantitative point-of-view we
will consider two metrics, namely, the \emph{variation range} and the
\emph{average normalised absolute dispersion error}. For a given
objective, the variation range represents the ratio between the
maximum and the minimum value observed for that objective. A
comparison between the variation range for different benchmarks
between a PS and a GA exploration for both power dissipation and
execution time is shown in Fig.~\ref{fig:dispersion}(a) and (b)
respectively. As it can be observed, the PS exploration provides solutions which fall on a
range that is, on average, wider than that provided by a
MOGA exploration for power dissipation and execution time, respectively.
\begin{table}
  \centering
  \begin{tabular}{c}
    \includegraphics[width=0.7\textwidth]{pictures/variation_power.eps} \\
    (a) \\
    \includegraphics[width=0.7\textwidth]{pictures/variation_etime.eps} \\
    (b) \\
    \includegraphics[width=0.7\textwidth]{pictures/dispersion.eps} \\
    (c) 
  \end{tabular}
  \caption{(a)(b) Variation range for different benchmarks between PS and GA
  and average dispersion absolute dispersion (c)}
  \label{fig:dispersion}
\end{table}

The average normalised absolute dispersion error measures
the average absolute difference between the distribution of points in
the objective space and an ideal distribution in which the points are
uniformly distributed over the objective space. Formally, let $O$ be
the image, in the objective space, of the configurations visited by
the design space exploration. The generic element of $O$ (\ie, a
solution) is a pair $(p,t)$ where $p$ and $t$ are the average power
and execution time, respectively. The two-dimensional objective space
is then partitioned by a $M_x \times M_y$ mesh. For each tile $T_i,
\ i=1, 2, \ldots, M_xM_y$ of the mesh, let $N_i$ be the number
of points in $O$ which fall in $T_i$. The average absolute error, $E_i$, for
$T_i$ is the absolute value of the difference between $N_i$ and the
ideal number of solutions, $\overline{N}$, which should fall in $T_i$
in case of uniform distribution. Such $\overline{N}$ can be simply
computed as the ratio between the cardinality of $O$ and the number of
tiles. Thus,
\[ E_i = |N_i - \overline{N}|, \]
where $\overline{N} = |O| / (M_x M_y)$. The average
normalised absolute dispersion error ($ANADE$) is the average of $E_i$
normalised to the maximum absolute error $E_{max}$:
\[ ANADE = \frac{\sum_{i=1}^{M_xM_y} E_i/(M_xM_y)}{E_{max}}, \]
where $E_{max}$ can be computed as the average absolute error in the worst
case in which all the solutions fall in a single tile:
\[ E_{max} = \frac{(M_x M_y - 1) \overline{N} + |\overline{N} -
    |O|| }{M_x M_y}. \]

>>>>

As regards the class of benchmark being considered, a set applications representative of some common
frequently running code kernels has been selected. Table~\ref{tab:bench} shows the set of applications
along with a brief description.
% IEEE table style
%\begin{table}
%	\centering
%	\caption{Benchmarks}
%	\label{tab:bench}
%	\begin{tabular}{ll}
%	\hline
%	\multicolumn{1}{c}{Benchmark} & \multicolumn{1}{c}{Application} \\
%	\hline
%	Alloca\_test & Memory array allocation test \\
%	Bmm & Matrices multiplication and elements sum \\
%	Fir\_int & Finite impulse response \\
%	Mm & Floating point matrices multiplication \\
%	Sha256 & Cryptocurrency header Hashing \\
%	Wave & Wavefront computation \\
%	\hline
%	\end{tabular}
%\end{table}

% ACM table style
\begin{table}
	\tbl{Benchmarks\label{tab:bench}}{
	\begin{tabular}{ll}
	\hline
	\multicolumn{1}{c}{Benchmark} & \multicolumn{1}{c}{Application} \\
	\hline
	Alloca\_test & Memory array allocation test \\
	Bmm & Matrices multiplication and elements sum \\
	Fir\_int & Finite impulse response \\
	Mm & Floating point matrices multiplication \\
	Sha256 & Cryptocurrency header Hashing \\
	Wave & Wavefront computation \\
	\hline
	\end{tabular}}
\end{table}
\subsection{Evaluation Metrics}

@@

% IEEE table style
%\begin{table}
%  \centering
%  \begin{tabular}{ccc}
%    \includegraphics[width=0.30\textwidth]{pictures/alloca_100.eps} &
%    \includegraphics[width=0.30\textwidth]{pictures/bmm_100.eps} & 
%    \includegraphics[width=0.30\textwidth]{pictures/fir_int100.eps} \\
%    \includegraphics[width=0.30\textwidth]{pictures/mm_100.eps} &
%    \includegraphics[width=0.30\textwidth]{pictures/sha_100.eps} &
%    \includegraphics[width=0.30\textwidth]{pictures/wave_100.eps} 
%  \end{tabular}
%  \caption{Pareto fronts found by PS and GA for a fixed budget of 100 configurations.}
%  \label{fig:pareto_fronts_100}
%\end{table}
%
%\begin{table}
%  \centering
%  \begin{tabular}{ccc}
%    \includegraphics[width=0.30\textwidth]{pictures/alloca_500.eps} &
%    \includegraphics[width=0.30\textwidth]{pictures/bmm_500.eps} & 
%    \includegraphics[width=0.30\textwidth]{pictures/fir_int500.eps} \\
%    \includegraphics[width=0.30\textwidth]{pictures/mm_500.eps} &
%    \includegraphics[width=0.30\textwidth]{pictures/sha_500.eps} &
%    \includegraphics[width=0.30\textwidth]{pictures/wave_500.eps} 
%  \end{tabular}
%  \caption{Pareto fronts found by PS and GA for a fixed budget of 500 configurations.}
%  \label{fig:pareto_fronts_500}
%\end{table}

\begin{table}
  \tbl{Pareto fronts found by PS and GA for a fixed budget of 100 configurations.\label{fig:pareto_fronts_100}}{
  \begin{tabular}{ccc}
    \includegraphics[width=0.30\textwidth]{pictures/alloca_100.eps} &
    \includegraphics[width=0.30\textwidth]{pictures/bmm_100.eps} & 
    \includegraphics[width=0.30\textwidth]{pictures/fir_int100.eps} \\
    \includegraphics[width=0.30\textwidth]{pictures/mm_100.eps} &
    \includegraphics[width=0.30\textwidth]{pictures/sha_100.eps} &
    \includegraphics[width=0.30\textwidth]{pictures/wave_100.eps} 
  \end{tabular}}

\end{table}

\begin{table}
  \tbl{Pareto fronts found by PS and GA for a fixed budget of 500
  configurations.\label{fig:pareto_fronts_500}}{
  \begin{tabular}{ccc}
    \includegraphics[width=0.30\textwidth]{pictures/alloca_500.eps} &
    \includegraphics[width=0.30\textwidth]{pictures/bmm_500.eps} & 
    \includegraphics[width=0.30\textwidth]{pictures/fir_int500.eps} \\
    \includegraphics[width=0.30\textwidth]{pictures/mm_500.eps} &
    \includegraphics[width=0.30\textwidth]{pictures/sha_500.eps} &
    \includegraphics[width=0.30\textwidth]{pictures/wave_500.eps} 
  \end{tabular}}
\end{table}
Low budget Pareto sets are shown in
Figure~\ref{fig:pareto_fronts_100}. With a so small number of
configuration being investigated the two sets are often overlapping,
being also strongly depended upon the application being considered.
These results are obtained in about 5 eras, and the main aim was
demonstrating that with a so limited budget PS still can compete with
genetic approach while even discovering very different Pareto fronts.
Continuing with a more extended simulation budget as depicted in
Figure~\ref{fig:pareto_fronts_500}, we can qualitatively observe that
even if PS does not strictly outperforms genetic in terms of dominance, it shows in
some cases largely Pareto in terms of extension. We can intuitively
relate this to the ``novelty-based'' score system, so that instead of
focusing on getting better points inside a given range it tries to
enlarge the range itself, resulting in less dense but more extended
Paretos. Of course we cannot classify this different behaviour as
``better'' or ``worse'' as compared to the genetic one, it depends on
various design factors such as desired results granularity, prefixed objective
range constraints, estimated error of the models used etc~\ldots.

In order to evaluate the Paretos from a quantitative point-of-view we
will consider two metrics, namely, the \emph{variation range} and the
\emph{average normalised absolute dispersion error}. For a given
objective, the variation range represents the ratio between the
maximum and the minimum value observed for that objective. A
comparison between the variation range for different benchmarks
between a PS and a GA exploration for both power dissipation and
execution time is shown in Fig.~\ref{fig:dispersion}(a) and (b)
respectively. As it can be observed, the PS exploration provides solutions which fall on a
range that is, on average, wider than that provided by a
MOGA exploration for power dissipation and execution time, respectively.
%\begin{table}
%  \centering
%  \begin{tabular}{c}
%    \includegraphics[width=0.7\textwidth]{pictures/variation_power.eps} \\
%    (a) \\
%    \includegraphics[width=0.7\textwidth]{pictures/variation_etime.eps} \\
%    (b) \\
%    \includegraphics[width=0.7\textwidth]{pictures/dispersion.eps} \\
%    (c) 
%  \end{tabular}
%  \caption{(a)(b) Variation range for different benchmarks between PS and GA
%  and average dispersion absolute dispersion (c)}
%  \label{fig:dispersion}
%\end{table}

\begin{table}
  \tbl{(a)(b) Variation range for different benchmarks between PS and GA
  and average dispersion absolute dispersion (c)
  \label{fig:dispersion}}{
  \begin{tabular}{c}
    \includegraphics[width=0.7\textwidth]{pictures/variation_power.eps} \\
    (a) \\
    \includegraphics[width=0.7\textwidth]{pictures/variation_etime.eps} \\
    (b) \\
    \includegraphics[width=0.7\textwidth]{pictures/dispersion.eps} \\
    (c) 
  \end{tabular}}
\end{table}
The average normalised absolute dispersion error measures
the average absolute difference between the distribution of points in
the objective space and an ideal distribution in which the points are
uniformly distributed over the objective space. Formally, let $O$ be
the image, in the objective space, of the configurations visited by
the design space exploration. The generic element of $O$ (\ie, a
solution) is a pair $(p,t)$ where $p$ and $t$ are the average power
and execution time, respectively. The two-dimensional objective space
is then partitioned by a $M_x \times M_y$ mesh. For each tile $T_i,
\ i=1, 2, \ldots, M_xM_y$ of the mesh, let $N_i$ be the number
of points in $O$ which fall in $T_i$. The average absolute error, $E_i$, for
$T_i$ is the absolute value of the difference between $N_i$ and the
ideal number of solutions, $\overline{N}$, which should fall in $T_i$
in case of uniform distribution. Such $\overline{N}$ can be simply
computed as the ratio between the cardinality of $O$ and the number of
tiles. Thus,
\[ E_i = |N_i - \overline{N}|, \]
where $\overline{N} = |O| / (M_x M_y)$. The average
normalised absolute dispersion error ($ANADE$) is the average of $E_i$
normalised to the maximum absolute error $E_{max}$:
\[ ANADE = \frac{\sum_{i=1}^{M_xM_y} E_i/(M_xM_y)}{E_{max}}, \]
where $E_{max}$ can be computed as the average absolute error in the worst
case in which all the solutions fall in a single tile:
\[ E_{max} = \frac{(M_x M_y - 1) \overline{N} + |\overline{N} -
    |O|| }{M_x M_y}. \]
